---
title: "IDS investigation worksheet"
author: "by Kate Supremacy: Connor, Kate, Ruth, Kate & Torgua"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
# Add any other libraries here



```


```{r load-data}
# load your data 
race_results <- read_csv("data/formula_e_race_results.csv") 
race_results %>%
  is.na() %>%
  colSums() # found all na values, these occur in laps and time_retired

race_results <- race_results %>%
  filter(!is.na(laps)) # removed oliver turvey from the dataset because he withdrew from the 2018 nyc race

race_results <- race_results %>%
  arrange(race_date, rank_num) # sort the results by data and driver ranking

race_results %>%
  filter(is.na(time_retired)) # display races where time_retired is NA

race_results <- race_results %>%
  separate(race_name, into = c("year", "race_location"), sep = 4, fill = "right") %>%
  select(-year)
# remove the race year just leaving location

race_results <- race_results %>%
  separate(race_location, into = c("race_loc", "loc_series"), sep = "race", fill = "right")
# added series for when multiple races in same location in a year

race_results$loc_series <- replace_na(race_results$loc_series, "1")
#set NAs created by above command to 1 (as only 1 race series in the year)

```



```{r}
points_breakdown <- race_results %>% 
  select(,driver,points,pts_rank,pts_pole,pts_bonus)

points_breakdown <- points_breakdown %>% 
  group_by(driver) %>% 
  summarize(points = sum(points),
            pts_rank = sum(pts_rank),
            pts_pole = sum(pts_pole),
            pts_bonus = sum(pts_bonus)
            ) %>% 
  arrange(desc(points))    #finds total points scored by each driver and breaks down the types of points too

summary(points_breakdown) #Mean total points is 104, normal points 99.6, pole points 2.958 and bonus points is 1.458

points_breakdown <- points_breakdown %>% 
 mutate(
   class = case_when(
    points > 700 ~ "A",
    points > 400 & points <= 700 ~ "B",
    points > 250 & points <= 400 ~ "C",
    points > 100 & points <= 250 ~ "D",
    points > 25 & points <= 100 ~ "E",
    points > 10 & points <= 25 ~ "F",
    points > 0 & points <= 10 ~ "G",
    points == 0 ~ "H"
   ))  #implements classes, it was the easiest way I could think of to split the 72 different drivers into groups based on the points they've scored over the seasons.
  
class_n <- points_breakdown %>% 
  count(class)

class_legend <- c(
  "A: 700 + points",
  "B: 401 - 700 points",
  "C: 251 - 400 points",
  "D: 101 - 250 points",
  "E: 26 - 100 points",
  "F: 11 - 25 points",
  "G: 1 - 10 points",
  "H: 0 points"
) #is displayed as the legend at the side of the graph
  
ggplot(class_n,aes(x = class, y = n, fill = class)) +
  geom_bar(stat = "identity") +
  scale_fill_manual( 
    values = c("A" = "gold", "B" = "grey", "C" = "orange", "D" = "red",
               "E" = "green", "F" = "blue", "G" = "purple", "H" = "black"), #Colours can be changed was just the first ones that came to mind.
    labels = class_legend)+
   labs(
    title = "Number of Drivers per Class",
    x = "Class",
    y = "Number of Drivers",
    fill = "Driver Class (Points scored over all 7 seasons)"
  )
```

```{r}
#How many races has each driver been in? Which driver has been in the most races in the data set?
race_results %>%
  count(driver) %>%
  arrange(desc(n))
#Lucas di Grassi and Sam Bird

#Which driver has ranked first the most?
race_results %>%
  filter(pts_rank == 25) %>%
  count(driver) %>%
  arrange(desc(n))
#Sébastien Buemi

#Which driver has come last the most?
race_results %>%
  filter(pts_rank == 1) %>%
  count(driver) %>%
  arrange(desc(n))
#Stéphane Sarrazin

#Which team group has ranked first the most?
race_results %>%
  filter(pts_rank == 25) %>%
  count(team_group) %>%
  arrange(desc(n))
#Nissan-Renault

#How many races does the data set include? (With the assumption every race has a winner.)
race_results %>%
  filter(pts_rank == 25) %>%
  count()
#71 Which would align with the frequency tables above.


#How often does pole position covert into a win?
race_results %>%
  filter(points >= 28) %>% 
  count()
#In 26 races pole position has converted to a win.
```

```{r}
#how many races have taken place each month over the course of all the seasons?
race_results %>%
  ggplot(aes(y = fct_rev(month(race_date, label = TRUE, abbr=FALSE)))) +
  geom_bar(aes(fill = month(race_date) != 7)) +
  labs(
    title = "Number of races taking place each month", 
    y = "Month", 
    x = "Number of races",
  ) +
  scale_fill_manual(values = c("red", "grey"), guide= "none")


```

